---
title: "Large Sample Theory Project"
author: "Jieqi Tu"
date: "11/5/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggpubr)
```

The algorithm of this project:

* Draw n (n = 10, 20 and 30) samples from $Gamma(k=1.67, \theta=49.98)$.
* Compute $\bar x$ and $s$.
* Repeat 1000 times. Then we will get $(\bar x_1, s_1)，(\bar x_2, s_2)，…,(\bar x_{1000}, s_{1000})$.
* Compute $\bar{\bar{x}}=\frac{\sum\bar{x_i}}{1000}$, $\bar{s}=[\frac{\sum{(\bar{x_i}-\bar{\bar{x}}})^2}{1000}]^{\frac{1}{2}}$
* Compute $z=\frac{\bar{x_i}-\bar{\bar{x}}}{\bar s/\sqrt{n}}$.
* Draw density function based on $z_i, i=1, 2, …,1000$.
* Compute $\bar{\bar{s_1^2}}=\frac{\sum{(\bar{x_i}-\bar{\bar{x}})^2}}{1000}$ and $\bar{\bar{s_2^2}}=\frac{\sum{s_i^2}}{1000}$

In this project, we want to compare $\bar{\bar{s_1^2}}$, $\bar{\bar{s_2^2}}$ and $k\theta^2$
#### n = 10
```{r draw samples n10}
set.seed(24)
k = 1.67
n = 10
theta = 49.98
x_bar = numeric(1000)
s = numeric(1000)
for (i in 1:1000) {
  sample = rgamma(n, shape = k, scale = theta)
  x_bar[i] = mean(sample)
  s[i] = sd(sample)
}
mean(x_bar)
sd(x_bar)
z = (x_bar - mean(x_bar))/(sd(x_bar)/sqrt(n))
df = data.frame(z = z)
```

```{r density function n10}
density1 = ggplot(df, aes(x = z)) + geom_density() + theme_bw()
density1
```

```{r compute s1 and s2 n10}
s1_1 = (sd(x_bar))^2 * n; s1_1
s2_1 = sum(s^2)/1000; s2_1
```

#### n = 20
```{r draw samples n20}
set.seed(24)
n = 20
x_bar2 = numeric(1000)
s2 = numeric(1000)
for (i in 1:1000) {
  sample = rgamma(n, shape = k, scale = theta)
  x_bar2[i] = mean(sample)
  s2[i] = sd(sample)
}
mean(x_bar2)
sd(x_bar2)
z2 = (x_bar2 - mean(x_bar2))/(sd(x_bar2)/sqrt(n))
df2 = data.frame(z = z2)
```

```{r density function n20}
density2 = ggplot(df2, aes(x = z)) + geom_density() + theme_bw()
density2
```

```{r compute s1 s2 n20}
s1_2 = (sd(x_bar2))^2 * n; s1_2
s2_2 = sum(s2^2)/1000; s2_2
```

#### n = 30
```{r draw samples n30}
set.seed(24)
n = 30
x_bar3 = numeric(1000)
s3 = numeric(1000)
for (i in 1:1000) {
  sample = rgamma(n, shape = k, scale = theta)
  x_bar3[i] = mean(sample)
  s3[i] = sd(sample)
}
mean(x_bar3)
sd(x_bar3)
z3 = (x_bar3 - mean(x_bar3))/(sd(x_bar3)/sqrt(n))
df3 = data.frame(z = z3)
```

```{r density function n30}
density3 = ggplot(df3, aes(x = z)) + geom_density() + theme_bw()
density3
ggarrange(density1, density2, density3,
          labels = c(1, 2, 3),
          ncol = 3, nrow = 1)
```

From the plot panel, we could see that, although they are all bell-shaped and symmetric distributed, the peak density became lower and lower as the sample size increases.
```{r compute s1 s2 n30}
s1_3 = (sd(x_bar3))^2 * n; s1_3
s2_3 = sum(s3^2)/1000; s2_3
```

#### Compare 3 scenarios
```{r compare three}
s1 = c(s1_1, s1_2, s1_3)
s2 = c(s2_1, s2_2, s2_3)
compare = data.frame(s1, s2)
true = k*theta^2
true
compare$'s1-true' = s1 - true
compare$'s2-true' = s2 - true
compare
```

